Transformation of forecasts for evaluating predictive performance in an epidemiological context

### 2023-05-07 main.tex
update manuscript in response to reviewers

Forecast evaluation plays an essential role in the development cycle of predictive epidemic models and can inform their use for public health decision-making. Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which are both measures of the absolute distance between the forecast distribution and the observation. They are commonly applied directly to predicted and observed incidence counts, but it can be questioned whether this is the optimal procedure for comparing models given the exponential nature of epidemic processes and the several orders of magnitude that observed values can span over space and time. In this paper, we argue that transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results. We motivate the procedure threefold using the CRPS on log-transformed counts as an example: Firstly, it can be interpreted as a probabilistic version of a relative error. Secondly, it reflects how well models predicted the time-varying epidemic growth rate. And lastly, using arguments on variance-stabilizing transformations, it can be shown that under the assumption of a quadratic mean-variance relationship, the logarithmic transformation leads to expected CRPS values which are independent of the order of magnitude of the predicted quantity. Applying the log transformation to data and forecasts from the European COVID-19 Forecast Hub, we find that it changes model rankings regardless of stratification by forecast date, location or target types. Situations in which models missed the beginning of upward swings are more strongly emphasized while failing to predict a downturn following a peak is less severely penalized. We conclude that appropriate transformations, of which the natural logarithm is only one particularly attractive option, should be considered when assessing the performance of different models in the context of infectious disease incidence. 
\end{abstract}

@@

Probabilistic forecasts are usually evaluated using so-called proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data. Proper scoring rules are constructed such that forecasters (anyone or anything that issues a forecast) are incentivised to report their true belief about the future. Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the Weighted Interval Score \citep{bracherEvaluatingEpidemicForecasts2021}. 
The CRPS measures the distance of the predictive distribution to the observed data as 
\begin{linenomath*}
\begin{equation}
    \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - \boldsymbol{1}(x \geq y) \right)^2 dx,
\end{equation}    
\end{linenomath*}
where $y$ is the true observed value, $F$ is the cumulative distribution function (CDF) of the predictive distribution, and $\boldsymbol{1}()$ is the indicator function. The CRPS can be understood as a generalisation of the absolute error to predictive distributions, and interpreted on the natural scale of the data. The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in so-called COVID-19 Forecast Hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}. The WIS is defined as 
\begin{linenomath*}
\begin{equation}
    \text{WIS}(F, y) = \frac{1}{K} \times \sum_{k = 1}^{K} 2 \times \left[ \boldsymbol{1}(y \leq q_{\tau_k}) - \tau_k \right] \times ( q_{\tau_k} - y), 
\end{equation}
\end{linenomath*}
where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, $K$ is the number of predictive quantiles provided and $\boldsymbol{1}()$ is the indicator function. The WIS can be decomposed into three components, dispersion, overprediction, underprediction, which reflect the width of the forecast and whether it was centred above or below the observed value. We show an alternative definition based on central prediction intervals in Section \ref{sec:alternative-wis} which illustrates this decomposition.

The dynamics of infectious processes are often described by the complementary concepts of the reproduction number $R$ \citep{gosticPracticalConsiderationsMeasuring2020} and growth rate $r$ \citep{wallingaHowGenerationIntervals2007}, where $R$ describes the strength and $r$ the speed of epidemic growth \citep{dushoffSpeedStrengthEpidemic2021}.
In the absence of changes in immunity, behaviour or other factors that may affect the intensity of transmission, the reproduction number would be expected to remain approximately constant.
In that case, the number of new infections in the population grows exponentially in time.
This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellisChallengesControlCOVID192021}.

@@

In many fields, it is common practice to forecast transformed quantities (see e.g. \cite{taylorEvaluatingVolatilityInterval1999} in finance, \cite{mayrLogLevelVAR2015} in macroeconomics, \cite{loweStochasticRainfallrunoffForecasting2014} in hydrology or \cite{fuglstadDoesNonstationarySpatial2015} in meteorology). While the goal of the transformations is usually to improve the accuracy of the predictions, they can also be used to enhance and complement the evaluation process. 
In this paper, we argue that the aforementioned issues with evaluating epidemic forecasts based on measures of absolute error on the natural scale can be addressed by transforming the forecasts and observations prior to scoring using some strictly monotonic transformation. Strictly monotonic transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while preserving the propriety of the score. Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects of the forecasts that forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}).

@@

When evaluating models across sets of forecasting tasks, it may be desirable for each target to have a similar impact on the overall results. In disease incidence forecasting, this is not the case when using the CRPS on the natural scale, as the latter typically scales with the order of magnitude of the quantity to be predicted. Average scores are then dominated by the results achieved for targets with high expected outcomes.

@@

To strengthen our intuition on how transforming outcomes prior to applying the CRPS shifts the emphasis between targets with high and low expected outcomes, Figure \ref{fig:SIM-wis-state-size-mean} shows the expected CRPS of ideal forecasters under different mean-variance relationships and transformations. We consider a Poisson distribution where $\sigma^2 = \mu$, a negative binomial distribution with size parameter $\theta = 10$ and thus $\sigma^2 = \mu + \mu^2/10$, and a normal distribution with constant variance. We see that when applying the CRPS on the natural scale, the expected CRPS grows with the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster). The expected CRPS is constant only for the distribution with constant variance, and grows in $\mu$ for the other two. When applying a log-transformation first, the expected CRPS is almost independent of $\mu$ for the negative binomial distribution and large $\mu$, while smaller targets have higher expected CRPS in case of the Poisson distribution and the normal distribution with constant variance. When applying a square-root-transformation before the CRPS, the expected CRPS is independent of the mean for the Poisson-distribution, but not for the other two (with a positive relationship in the normal case and a negative one for the negative binomial). As can be seen in Figures \ref{fig:SIM-wis-state-size-mean} and \ref{fig:score-approx}, the approximations presented above work quite well for our simulated example.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/SIM-mean-state-size.png}
    \caption{Expected CRPS scores as a function of the mean and variance of the forecast quantity. We computed expected CRPS values  for three different distributions, assuming an ideal forecaster with predictive distribution equal to the data-generating distribution. 
    These expected CRPS values were computed for different predictive means based on 10,000 samples each and are represented by dots. Solid lines show the corresponding approximation of the expected CRPS based on an assumed normal distribution as discussed in section \ref{sec:methods:vst}. Figure \ref{fig:score-approx} shows the quality of the approximation in more detail. 
    The first distribution (red) is a truncated normal distribution with constant variance (we chose $\sigma = 1$ in order to only obtain positive samples). The second (green) is a negative binomial distribution with variance $\theta = 10$ and variance $\sigma^2 = \mu + 0.1\mu^2$. The third (blue) is a
    Poisson distribution with $\sigma^2 = \mu$. To make the scores for the different distributions comparable, scores were normalised to one, meaning that the mean score for every distribution (red, green, blue) is one. 
    A: Normalised expected CRPS for ideal forecasts with increasing means for three distribution with different relationships between mean and variance. Expected CRPS was computed on the natural scale (left), after applying a square-root transformation (middle), and after adding one and applying a log-transformation to the data (right). B: A but with x axis on the log scale.}
    \label{fig:SIM-wis-state-size-mean}
\end{figure}

@@

Transformations that are strictly monotonic are permissible in the sense that they maintain the propriety of the score. This is because even though rankings of models may change forecasts will in expectation still minimise their score if they report a predictive distribution that is equal to the data-generating distribution. This condition holds for both the log and square root transformations, as well as many others. However, the order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee propriety. In the case of log transforms, taking the logarithm of the scores, rather than scoring the log-transformed forecasts and data, results in an improper score. This is because taking the logarithm of the CRPS (or WIS) results in a score that does not penalise outliers enough and therefore incentivises overconfident predictions. We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can easily be seen that overconfident models perform best in terms of the log WIS.

In practice, one issue with the log transform is that they are not readily applicable to negative numbers or zero values, which need to be removed or otherwise handled. 
One common approach to deal with zeros is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm \citep{bellegoDealingLogsZeros2022}. This represents a strictly monotonic transformation and therefore preserves the propriety of the resulting score. The choice of the quantity to add does however influence scores and rankings, as measures of relative errors shrink when adding a constant $a$ to the forecast and the observation. We illustrate this in Figure \ref{fig:illustration-effect-log-offset}. As a rule of thumb, if if $x > 5a$, the difference between $\log{(x + a)}$ and $\log{(x)}$ is small, and it becomes negligible if $x > 50a$. Choosing a suitable offset $a$ balances two competing concerns: on the one hand, choosing a small $a$ makes sure that the transformation is as close to a natural logarithm as possible and scores can be interpreted as outlined above. On the other hand, choosing a larger $a$ can help stabilise scores for forecasts and observations close to zero, avoiding giving excessive weight to forecasts for small quantities (see Figure \ref{fig:HUB-log-different-offsets}).

@@

The forecasts used for the purpose of this illustration are forecasts submitted between the 8th of March 2021 and the 5th of December 2022 for reported cases and deaths from COVID-19. See \cite{sherrattPredictivePerformanceMultimodel2022} for a more thorough description of the data. We filtered all forecasts submitted to the Hub to only include models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 46 forecast dates (see Figure \ref{fig:HUB-num-avail-models}). We removed all observations marked as data anomalies by the European Forecast Hub \citep{sherrattPredictivePerformanceMultimodel2022} as well as all remaining negative observed values. In addition, we filtered out erroneous forecasts defined by any of the conditions listed in Table \ref{tab:erroneous}. Those forecasts were removed in order to be better able to illustrate the effects of the log-transformation on scores and eliminating distortions caused by outlier forecasters. All predictive quantiles were truncated at 0. We applied the log-transformation after adding a constant $a = 1$ to all predictions and observed values. The choice of $a = 1$ in part reflects convention, but also represents a suitable choice as it avoids giving excessive weight to forecasts close to zero, while at the same time ensuring that scores for observations $> 5$ can be interpreted reasonably. The analysis was conducted in \texttt{R} \citep{R}, using the \texttt{scoringutils} package \citep{bosseEvaluatingForecastsScoringutils2022} for forecast evaluation. All code is available on GitHub 
(\url{https://github.com/epiforecasts/transformation-forecast-evaluation}). Where not otherwise stated, we report results for a two-week-ahead forecast horizon.

@@

When comparing examples of forecasts on the natural scale with those on the log scale (see Figures \ref{fig:HUB-model-comparison-ensemble}, \ref{fig:HUB-model-comparison-baseline}, \ref{fig:HUB-model-comparison-epinow}) a few interesting patterns emerge. Missing the peak, i.e. predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts in May in Figure \ref{fig:HUB-model-comparison-ensemble}A, B). On the log scale, these have less of an influence, as errors are smaller in relative terms (see \ref{fig:HUB-model-comparison-ensemble}C, D). Conversely, failure to predict an upswing while numbers are still low, is less severely punished on the natural scale (see forecasts in July in Figure \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low. On the log scale, missing lower inflection points tends to lead to more severe penalties (see Figure \ref{fig:HUB-model-comparison-ensemble}C, D)). One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Figure \ref{fig:HUB-model-comparison-ensemble}E, F). On the log scale, scores do not exhibit this behaviour and rather increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/HUB-mean-obs-location.png}
    \caption{Observations and scores across locations and forecast horizons for the European COVID-19 Forecast Hub data. Locations are sorted according to the mean observed value in that location. 
    A: Average (across all time points) of observed cases and deaths for different locations. B: Corresponding boxplot (y-axis on log-scale) of all cases and deaths. C: Scores for two-week-ahead forecasts from the EuroCOVIDhub-ensemble (averaged across all forecast dates) for different locations, evaluated on the natural as well as the logarithmic scale. D: Corresponding boxplots of all individual scores for two-week-ahead predictions. E: Boxplots for the relative change of scores for the EuroCOVIDhub-ensemble across forecast horizons. For any given forecast date and location, forecasts were made for four different forecast horizons, resulting in four scores. All scores were divided by the score for forecast horizon one. To enhance interpretability, the range of visible relative changes was restricted to [0.1, 10].}
    \label{fig:HUB-mean-locations}
\end{figure}

@@

To check the relationship after the transformation, we ran the regressions
\begin{equation}
    \text{WIS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},
\end{equation}
where $F_{\log}$ is the predictive distribution for $\log(y)$, and
\begin{equation}
    \text{WIS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{\sqrt{\ }} + \beta_{\sqrt{\ }} \cdot \sqrt{\text{median}(F)},
\end{equation} 
where $F_{\sqrt{\ }}$ is the predictive distribution on the square-root scale. A value of $\beta_{\log} = 0$ (or $\beta_{\sqrt{\ }} = 0$, respectively) would imply that scores are independent of the median prediction after the transformation. A value smaller (larger) than 0 would imply that smaller (larger) targets lead to higher scores. As can be seen from Table \ref{tab:HUB-regression}, the results indeed indicate that small targets lead to larger average WIS when using the log transform ($\beta_{\log} < 0$), while the opposite is true for the square-root transform ($\beta_{\sqrt{\ }} > 0$). The results of the three regressions are also displayed in Figure \ref{fig:HUB-regression}. In this empirical example, the log transformation thus helps (albeit not perfectly), to stabilise WIS values, and it does so more successfully than the square-root transformation. As can be seen from Figure \ref{fig:HUB-regression}, the expected WIS scores for case targets with medians of 10 and 100,000 differ by more then a factor of ten for the square root transformation, but only a factor of around 2 for the logarithm.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/HUB-correlations.png}
    \caption{Correlations of rankings on the natural and logarithmic scale. A: Average Spearman rank correlation of scores for individual forecasts from all models. For every individual target (defined by a combination of forecast date, target type, horizon, location), one score was obtained per model. Then, the rank correlation was computed between the scores for all models on the natural scale vs. on the log scale. All rank correlations were averaged across locations, and target types and stratified by horizon and target type. B: Correlation between relative skill scores. For every forecast horizon and target type, a separate relative skill score was computed per model using pairwise comparisons. The plot shows the correlation between the relative skill scores on the natural vs. on the log scale.}
    \label{fig:HUB-cors}
\end{figure}

For \textit{individual} forecasts, rankings between models for single forecasts are mostly preserved, with differences increasing across forecast horizons (see Figure \ref{fig:HUB-cors}A). When evaluating performance \textit{averaged across} different forecasts and forecast targets, relative skill scores of the models change considerably (Figure \ref{fig:HUB-cors}B). The correlation between relative skill scores also decreases noticeably with increasing forecast horizon.

>>>>

Forecast evaluation is essential for the development of predictive epidemic models and can inform their use for public health decision-making. Common scores to evaluate epidemiological forecasts are the Continuous Ranked Probability Score (CRPS) and the Weighted Interval Score (WIS), which are both measures of the absolute distance between the forecast distribution and the observation. However, applying these scores directly to predicted and observed incidence counts may not be optimal due to the exponential nature of epidemic processes and the varying magnitudes of observed values across space and time. In this paper, we argue that transforming counts before applying scores such as the CRPS or WIS can effectively mitigate these difficulties and yield epidemiologically meaningful and easily interpretable results. We motivate the procedure threefold using the CRPS on log-transformed values as an example: Firstly, it can be interpreted as a probabilistic version of a relative error. Secondly, it reflects how well models predicted the time-varying epidemic growth rate. And lastly, using arguments on variance-stabilizing transformations, it can be shown that under the assumption of a quadratic mean-variance relationship, the logarithmic transformation leads to expected CRPS values which are independent of the order of magnitude of the predicted quantity. Applying a transformation of log(x + 1) to data and forecasts from the European COVID-19 Forecast Hub, we find that it changes model rankings regardless of stratification by forecast date, location or target types. Situations in which models missed the beginning of upward swings are more strongly emphasised while failing to predict a downturn following a peak is less severely penalised when scoring transformed forecasts as opposed to untransformed ones. We conclude that appropriate transformations, of which the natural logarithm is only one particularly attractive option, should be considered when assessing the performance of different models in the context of infectious disease incidence.
\end{abstract}

@@

Probabilistic forecasts are usually evaluated using so-called proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data. Proper scoring rules are constructed such that they cannot be gamed, meaning that forecasters (anyone or anything that issues a forecast) are incentivised to report their true belief about the future. In expectation, a forecaster receives the best score if their forecast $F$ coincides with the data-generating distribution $G$, that is if $F = G$. Common proper scoring rules are the logarithmic score \citep{goodRationalDecisions1952} or the continuous ranked probability score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007}. The log score is the log-likelihood of the predictive density at the observed value and is scale-invariant under monotonous transformations. The log score does not suffer from the issues we address in this paper, but due to a number of reasons is not widely used in fields such as weather forecasting or epidemiology. For example, the log score becomes very unstable for outlier forecasts that assign low density to the observed outcome. It is also difficult to apply to forecasts of discrete quantities and doesn't work for forecasts in a quantile-based format. 
More common in epidemiology are the CRPS or its discrete equivalent, the ranked probability score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the weighted interval score ~\citep[WIS,][]{bracherEvaluatingEpidemicForecasts2021}. 
The CRPS measures the distance of the predictive distribution to the observed data as 
\begin{linenomath*}
\begin{equation}
    \text{CRPS}(F, y) = \int_{-\infty}^\infty \left( F(x) - \boldsymbol{1}(x \geq y) \right)^2 dx,
\end{equation}    
\end{linenomath*}
where $y$ is the true observed value, $F$ is the cumulative distribution function (CDF) of the predictive distribution, and $\boldsymbol{1}()$ is the indicator function. The CRPS can be understood as a generalisation of the absolute error to predictive distributions, and interpreted on the natural scale of the data. The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in so-called COVID-19 Forecast Hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}. The WIS is defined as 
\begin{linenomath*}
\begin{equation}
    \text{WIS}(F, y) = \frac{1}{K} \times \sum_{k = 1}^{K} 2 \times \left[ \boldsymbol{1}(y \leq q_{\tau_k}) - \tau_k \right] \times ( q_{\tau_k} - y), 
\end{equation}
\end{linenomath*}
where $q_{\tau}$ is the $\tau$ quantile of the forecast $F$, $y$ is the observed outcome, $K$ is the number of predictive quantiles provided and $\boldsymbol{1}()$ is the indicator function. The WIS can be decomposed into three components, dispersion, overprediction, underprediction, which reflect the width of the forecast and whether it was centred above or below the observed value. We show an alternative definition based on central prediction intervals in Section \ref{sec:alternative-wis} which illustrates this decomposition.

The dynamics of infectious processes are often described by the complementary concepts of the reproduction number $R$ \citep{gosticPracticalConsiderationsMeasuring2020} and growth rate $r$ \citep{wallingaHowGenerationIntervals2007}. The reproduction number $R$ describes the number of people each infected person is expected to infect in turn (irrespective of the time between infections) and therefore is a measure of the strength of epidemic growth. The growth rate $r$ describes the speed of epidemic growth \citep{dushoffSpeedStrengthEpidemic2021}.
In the absence of changes in population immunity, behaviour or other factors that may affect the intensity of transmission, the reproduction number would be expected to remain approximately constant.
In that case, the number of new infections in the population grows exponentially in time.
This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellisChallengesControlCOVID192021}.

@@

In many fields, it is common practice to forecast transformed quantities (see e.g. \cite{taylorEvaluatingVolatilityInterval1999} in finance, \cite{mayrLogLevelVAR2015} in macroeconomics, \cite{loweStochasticRainfallrunoffForecasting2014} in hydrology or \cite{fuglstadDoesNonstationarySpatial2015} in meteorology). While the goal of the transformations is usually to improve the accuracy of the predictions, they can also be used to enhance and complement the evaluation process. 
In this paper, we argue that the aforementioned issues with evaluating epidemic forecasts based on measures of absolute error on the natural scale can be addressed by transforming the forecasts and observations prior to scoring using some strictly monotonic transformation. Strictly monotonic transformations can shift the focus of the evaluation in a way that may be more appropriate for epidemiological forecasts, while guaranteeing that the score remains proper. Many different transformations may be appropriate and useful, depending on the exact context, the desired focus of the evaluation, and specific aspects of the forecasts that forecast consumers care most about (see a broader discussion in Section \ref{sec:discussion}).

@@

When evaluating models across sets of forecasting tasks, it may be desirable for each target to have a similar impact on the overall results. This is true if one assumes that forecasts for different targets provide similar amounts of information about how well a forecaster performs. Ideally, one would like the magnitude of the score to scale with the difficulty of the prediction tasks and the information that can be obtained about overall performance of the forecaster. In disease incidence forecasting, however, CRPS values on the natural scale typically scale with the order of magnitude of the quantity to be predicted. Average scores are then dominated by the results achieved for targets with high expected outcomes in a way that does not necessarily reflect the underlying predictive ability well.

@@

To strengthen our intuition on how transforming outcomes prior to applying the CRPS shifts the emphasis between targets with high and low expected outcomes, Figure \ref{fig:SIM-wis-state-size-mean} shows the expected CRPS of ideal forecasters under different mean-variance relationships and transformations. We consider a Poisson distribution where $\sigma^2 = \mu$, a negative binomial distribution with size parameter $\theta = 10$ and thus $\sigma^2 = \mu + \mu^2/10$, and a normal distribution with constant variance. We see that when applying the CRPS on the natural scale, the expected CRPS grows as the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster) increases. The expected CRPS is constant only for the distribution with constant variance, and grows in $\mu$ for the other two. When applying a log-transformation first, the expected CRPS is almost independent of $\mu$ for the negative binomial distribution and large $\mu$, while smaller targets have higher expected CRPS in case of the Poisson distribution and the normal distribution with constant variance. When applying a square-root-transformation before the CRPS, the expected CRPS is independent of the mean for the Poisson-distribution, but not for the other two (with a positive relationship in the normal case and a negative one for the negative binomial). As can be seen in Figures \ref{fig:SIM-wis-state-size-mean} and \ref{fig:score-approx}, the approximations presented above work quite well for our simulated example.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/SIM-mean-state-size.png}
    \caption{Expected CRPS scores as a function of the mean and variance of the forecast quantity. We computed expected CRPS values  for three different distributions, assuming an ideal forecaster with predictive distribution equal to the true underlying (data-generating) distribution. 
    These expected CRPS values were computed for different predictive means based on 10,000 samples each and are represented by dots. Solid lines show the corresponding approximation of the expected CRPS based on an assumed normal distribution as discussed in section \ref{sec:methods:vst}. Figure \ref{fig:score-approx} shows the quality of the approximation in more detail. 
    The first distribution (red) is a truncated normal distribution with constant variance (we chose $\sigma = 1$ in order to only obtain positive samples). The second (green) is a negative binomial distribution with variance $\theta = 10$ and variance $\sigma^2 = \mu + 0.1\mu^2$. The third (blue) is a
    Poisson distribution with $\sigma^2 = \mu$. To make the scores for the different distributions comparable, scores were normalised to one, meaning that the mean score for every distribution (red, green, blue) is one. 
    A: Normalised expected CRPS for ideal forecasts with increasing means for three distribution with different relationships between mean and variance. Expected CRPS was computed on the natural scale (left), after applying a square-root transformation (middle), and after adding one and applying a log-transformation to the data (right). B: A but with x axis on the log scale.}
    \label{fig:SIM-wis-state-size-mean}
\end{figure}

@@

Transformations that are strictly monotonic are permissible in the sense that resulting scores remain proper. This is because even though rankings of models may change forecasts will in expectation still minimise their score if they report a predictive distribution that is equal to the data-generating distribution. This condition holds for both the log and square root transformations, as well as many others. However, the order of the operations matters, and applying a transformation after scores have been computed generally does not guarantee that the score remains proper. In the case of log transforms, taking the logarithm of the scores, rather than scoring the log-transformed forecasts and data, results in an improper score. This is because taking the logarithm of the CRPS (or WIS) results in a score that does not penalise outliers enough and therefore incentivises overconfident predictions. We illustrate this point using simulated data in Figure \ref{fig:log-improper}, where it can easily be seen that overconfident models perform best in terms of the log WIS.

In practice, one issue with the log transform is that they are not readily applicable to negative numbers or zero values, which need to be removed or otherwise handled. 
One common approach to deal with zeros is to add a small quantity, such as 1, to all observations and predictions before taking the logarithm \citep{bellegoDealingLogsZeros2022}. This represents a strictly monotonic transformation and therefore guarantees that the resulting score remains proper. The choice of the quantity to add does however influence scores and rankings, as measures of relative errors shrink when adding a constant $a$ to the forecast and the observation. We illustrate this in Figure \ref{fig:illustration-effect-log-offset}. As a rule of thumb, if if $x > 5a$, the difference between $\log{(x + a)}$ and $\log{(x)}$ is small, and it becomes negligible if $x > 50a$. Choosing a suitable offset $a$ balances two competing concerns: on the one hand, choosing a small $a$ makes sure that the transformation is as close to a natural logarithm as possible and scores can be interpreted as outlined above. On the other hand, choosing a larger $a$ can help stabilise scores for forecasts and observations close to zero, avoiding giving excessive weight to forecasts for small quantities (see Figure \ref{fig:HUB-log-different-offsets}).

@@

The forecasts used for the purpose of this illustration are forecasts submitted between the 8th of March 2021 and the 5th of December 2022 for reported cases and deaths from COVID-19. Target dates range from the 13th of March 2021 to the 10th of December 2022, for a total of 92 weeks. See \cite{sherrattPredictivePerformanceMultimodel2022} for a more thorough description of the data. We filtered all forecasts submitted to the Hub to only include the seven models which have submitted forecasts for both deaths and cases for 4 horizons in 32 locations on at least 46 forecast dates (see Figure \ref{fig:HUB-num-avail-models}). We removed all observations marked as data anomalies by the European Forecast Hub \citep{sherrattPredictivePerformanceMultimodel2022} as well as all remaining negative observed values. These anomalies made up a relevant fraction of all observations. On average across locations, 12.1 out of 92 (13.2\%) observations were removed for cases and 12.4 out of 92 (13.5\%) for deaths. Figure \ref{fig:number-anomalies} displays the number of anomalies removed for each location. In addition, we filtered out erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table \ref{tab:erroneous}. Figure \ref{fig:erroneous-forecasts.png} shows the percentage of forecasts removed for each model. Those few (less than 0.2\%  of forecasts for each model) erroneous outlier forecasts had excessive influence on average scores and relative skill scores in a way that was not representative of normal model behaviour. We removed them here in order to better illustrate the effects of the log-transformation on scores that one would expect in a 'normal' scenario. In a regular forecast evaluation such erroneous forecasts should usually not be removed and would count towards overall model scores.

All predictive quantiles were truncated at 0. We applied the log-transformation after adding a constant $a = 1$ to all predictions and observed values. The choice of $a = 1$ in part reflects convention, but also represents a suitable choice as it avoids giving excessive weight to forecasts close to zero, while at the same time ensuring that scores for observations $> 5$ can be interpreted reasonably. The analysis was conducted in \texttt{R} \citep{R}, using the \texttt{scoringutils} package \citep{bosseEvaluatingForecastsScoringutils2022} for forecast evaluation. All code is available on GitHub 
(\url{https://github.com/epiforecasts/transformation-forecast-evaluation}). Where not otherwise stated, we report results for a two-week-ahead forecast horizon.

@@

When comparing examples of forecasts on the natural scale with those on the log scale (see Figures \ref{fig:HUB-model-comparison-ensemble}, \ref{fig:HUB-model-comparison-baseline}, \ref{fig:HUB-model-comparison-epinow}) a few interesting patterns emerge. Missing the peak, i.e. predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the natural scale (see forecasts during the peak in May 2022 in Figure \ref{fig:HUB-model-comparison-ensemble}A, B). On the log scale, these have less of an influence, as errors are smaller in relative terms (see \ref{fig:HUB-model-comparison-ensemble}C, D). Conversely, failure to predict an upswing while numbers are still low, is less severely penalised on the natural scale (see forecasts in July 2021 and to a lesser extent in July 2022 in Figure \ref{fig:HUB-model-comparison-ensemble} A, B), as overall absolute errors are low. On the log scale, missing lower inflection points tends to lead to more severe penalties (see Figure \ref{fig:HUB-model-comparison-ensemble}C, D)). One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Figure \ref{fig:HUB-model-comparison-ensemble}E, F). On the log scale, scores do not exhibit this behaviour and rather increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/HUB-mean-obs-location.png}
    \caption{Observations and scores across locations and forecast horizons for the European COVID-19 Forecast Hub data. Locations are sorted according to the mean observed value in that location. 
    A: Average (across all time points) of observed cases and deaths for different locations. B: Corresponding boxplot (y-axis on log-scale) of all cases and deaths. C: Scores for two-week-ahead forecasts from the EuroCOVIDhub-ensemble (averaged across all forecast dates) for different locations, evaluated on the natural scale as well as after transforming counts by adding one and applying the natural logarithm. D: Corresponding boxplots of all individual scores of the EuroCOVIDhub-ensemble for two-week-ahead predictions. E: Boxplots for the relative change of scores for the EuroCOVIDhub-ensemble across forecast horizons. For any given forecast date and location, forecasts were made for four different forecast horizons, resulting in four scores. All scores were divided by the score for forecast horizon one. To enhance interpretability, the range of visible relative changes in scores (relative to horizon = 1) was restricted to [0.1, 10].}
    \label{fig:HUB-mean-locations}
\end{figure}

@@

To check the relationship after the transformation, we ran the regressions
\begin{equation}
    \text{WIS}(F_{\log}, \log y) = \alpha_{\log} + \beta_{\log} \cdot \log{(\text{median}(F))},
\end{equation}
where $F_{\log}$ is the predictive distribution for $\log(y)$, and
\begin{equation}
    \text{WIS}(F_{\sqrt{\ }}, \sqrt{y}) = \alpha_{\sqrt{\ }} + \beta_{\sqrt{\ }} \cdot \sqrt{\text{median}(F)},
\end{equation} 
where $F_{\sqrt{\ }}$ is the predictive distribution on the square-root scale. A value of $\beta_{\log} = 0$ (or $\beta_{\sqrt{\ }} = 0$, respectively) would imply that scores are linearly independent of the median prediction after the transformation. A value smaller (larger) than 0 would imply that smaller (larger) targets lead to higher scores. As can be seen from Table \ref{tab:HUB-regression}, the results indeed indicate that small targets lead to larger average WIS when using the log transform ($\beta_{\log} < 0$), while the opposite is true for the square-root transform ($\beta_{\sqrt{\ }} > 0$). The results of the three regressions are also displayed in Figure \ref{fig:HUB-regression}. In this empirical example, the log transformation thus helps (albeit not perfectly), to stabilise WIS values, and it does so more successfully than the square-root transformation. As can be seen from Figure \ref{fig:HUB-regression}, the expected WIS scores for case targets with medians of 10 and 100,000 differ by more then a factor of ten for the square root transformation, but only a factor of around 2 for the logarithm.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/HUB-correlations.png}
    \caption{Correlations of rankings on the natural and logarithmic scale. A: Average Spearman rank correlation of scores for individual forecasts. For every individual target (defined by a combination of forecast date, target type, horizon, location), one score was obtained per model. Then, for every forecast target, the Spearman rank correlation was computed between scores on the natural scale and on the log scale for all the models that had made a forecast for that specific target. These individual rank correlations were then averaged across locations and time and are displayed stratified by horizon and target types, representing average accordance of model ranks for a single forecast target on the natural and on the log scale. B: Correlation between relative skill scores. For every forecast horizon and target type, a separate relative skill score was computed per model using pairwise comparisons, which is a measure of performance of a model relative to the others for a given horizon and target type that accounts for missing values. The plot shows the correlation between the relative skill scores on the natural vs. on the log scale, representing accordance of overall model performance as judged by scores on the natural and on the log scale.}
    \label{fig:HUB-cors}
\end{figure}
